---
title: "Prediction of Housing Values in Boston"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**File name:** "housing.proper.csv" (given in the problem set) has housing values in the suburbs of Boston  

- **506  observations** and **13 predictor variables** (independent variables or X or features)      
- Median price is a more accurate measure of market prices as it is not affected by outliers       

```{R}

library (tidyverse)  
library (ggplot2)  
library(corrplot)  

housing_data = read.csv("housing.proper.csv") #loading the dataset
housing_data_df= data.frame(housing_data)
colnames(housing_data_df) <- c("X1","X2","X3","X4","X5","X6","X7","X8",
                            "X9","X10","X11","X12","X13", "Y") #transforming column names
```
# Exploratory Data Analysis    

**Objective of EDA** : To choose important variables or features of this dataset
Visualisations are used the explain relationship between features and target variable    


**Split the data into 70% training set (model building data set) and 30% testing set**


```{r}
set.seed(1234) #set seed to ensure same random numbers are generated
require(caTools)
sample = sample.split(housing_data_df, SplitRatio = 0.7)
#Splits the dataset "housing_data" in the ratio mentioned. Marks rows as logical true and remaining as logical false.  

training_housing_data = subset( housing_data_df, sample == TRUE)  #creates training dataset with rows marked true
testing_housing_data = subset(housing_data_df, sample ==FALSE)
```
**Choosing features through Summary Statistics, Correlation Plot, Heat Map and Boxplot**  

EDA is performed on the complete dataset

```{r}

summary (housing_data_df) # to get statistic of each variable to detect outliers  

housing_data.cor = round(cor(housing_data_df), 3) #correlation between features and median housing values
housing_data.cor


heatmap(housing_data.cor, Rowv = NA, Colv = NA)

corrplot(cor(housing_data_df), method = "number", type = "upper", diag = FALSE)

housing_data.box = boxplot(housing_data_df, xlab= "Features", ylab ="Frequency") #boxplot to detect outliers   



```


